Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-05-11 Manuscript/proto-manuscript.tex
Small addition to the MS: simple formula for interaction energy with scaling.
Small edits in proto-MS

\begin{abstract}
% insert abstract here
Classical molecular dynamics simulations give detailed information about membrane structure and dynamics. However, there is still a room for improvements in current force fields – it is known from the literature, that the binding of ions, especially cations, to phopholipid membranes is overestimated in all classical models [1]. We suggest that the membrane-ion interactions can be corrected by including implicit electronic polarizability into the lipid models through the electronic continuum correction (ECC) [2], which was already applied to monovalent and divalent ions yielding models that feature correct ion pairing [3]. Using the electrometer concept [3, 4], our pilot simulations point out that our hypothesis is correct and ECC is indeed a missing important contribution in current classical lipid models. Moreover, the solid physical principles behind ECC are found not to hamper other relevant properties of a phospholipid bilayer. Poster presents a new lipid model, "ECClipids", that shows accurate affinity to sodium and calcium cations. 
This work will continue as an open collaboration project NMRlipids IV (http://nmrlipids.blogspot.fi).

@@

% Here I write pseudo-article statements that will make the main argument.
% Beautiful polished sentences will be formed only afte we agree on these basic things
\textbf{Introduction}

@@

\textbf{MDEC and current FFs}

@@

Succesful application: Works for cations \cite{Jungwirth2015,Kohagen..} -- motivation for its application to zwitterionic lipids like POPC.

Hypothesis: ECC helps in describing even zwitterionic molecules like POPC, will be demonstrated through headgroup oreder parametere response to cationic molecules.

@@

ECC and solvation $\Delta G$.(from \cite{Leontyev2015}: hydration $\Delta G$ can provide good structur and energetics, but will fail in good interactions -- already proved for cations)
In addition, $\Delta G_{hyd}$ -- the usual target for parametrization of small molecules in classical force fields -- is not the right target in the MDEC/ECC framework -- part of $\Delta G$ is already included in the polarization of electrons, and only the remining part of $\Delta G$ belongs to the polarization of the nuclei. 
Interestingly, the part of $\Delta G$ due to electrons is approximately $1/2$ of it, which is what I observe in my simulations -- if I assume that half of this energy is already accounted for by the water model (which is rather a MDEC model already), than I should recover approximately $3/4$ of the original interaction energy. 
When I write the formula for interaction energy for two particles at a distance $r=\sigma$ (approximate equilibrium position), I get
\begin{equation}
W(f_\sigma r,f_q q) = \kappa \frac{q_i q_j}{r} \frac{f_q^2}{f_\sigma}
\end{equation}

@@

So for $f_q=0.8$ I obtain $f_\sigma ~= \frac{4}{3} f_q^2 \,\,\, = 0.853$ -- which is very close to the value I use, $f_\sigma=0.875$. This is a crude derivation neglecting the true position of energy minimum and the LJ mixing rules.

>>>>

\begin{abstract}
% insert abstract here
Classical molecular dynamics simulations give detailed information about membrane structure and dynamics. However, there is still a room for improvements in current force fields – it is known from the literature, that the binding of ions, especially cations, to phopholipid membranes is overestimated in all classical models [1]. We suggest that the membrane-ion interactions can be corrected by including implicit electronic polarizability into the lipid models through the electronic continuum correction (ECC) [2], which was already applied to monovalent and divalent ions yielding models that feature correct ion pairing [3]. Using the electrometer concept [3, 4], our simulations point out that our hypothesis is correct and ECC is indeed a missing important contribution in current classical lipid models. Moreover, the solid physical principles behind ECC are found not to hamper other relevant properties of a phospholipid bilayer. 
The new lipid model, "ECClipids", shows accurate affinity to sodium and calcium cations. 
This work will continue as an open collaboration project NMRlipids IV (http://nmrlipids.blogspot.fi).

@@

% Here I write pseudo-article statements that will make the main argument.
% Beautiful polished sentences will be formed only afte we agree on these basic things
\textbf{General Introduction}

@@

\textbf{Specific introduction: MDEC and current FFs}

@@

Succesful application: Works for cations \cite{Jungwirth2015,Kohagen..}\todoii{REF}{Missing references} -- motivation for its application to zwitterionic lipids like POPC.

Hypothesis: ECC helps in describing even zwitterionic molecules like POPC, will be demonstrated through headgroup order parameter response to cationic molecules.

@@

ECC and solvation $\Delta G$.(from \cite{Leontyev2015}: hydration $\Delta G$ can provide good structure and energetics, but will fail in good interactions -- already proved for cations)
In addition, $\Delta G_{hyd}$ -- the usual target for parametrization of small molecules in classical force fields -- is not the right target in the MDEC/ECC framework -- part of $\Delta G$ is already included in the polarization of electrons, and only the remining part of $\Delta G$ belongs to the polarization of the nuclei. 
Interestingly, the part of $\Delta G$ due to electrons is approximately $1/2$ of it, which is what I observe in my simulations -- if I assume that half of this energy is already accounted for by the water model (which is rather a MDEC model already), than I should recover approximately $3/4$ of the original interaction energy. 
When I write the formula for interaction energy for two particles at a distance $r=\sigma$ (approximate equilibrium position), I get
\begin{equation}
W(f_\sigma r,f_q q) = \kappa \frac{q_i q_j}{r} \frac{f_q^2}{f_\sigma}
\end{equation}

@@

So for $f_q=0.8$ I obtain $f_\sigma ~= \frac{4}{3} f_q^2 \,\,\, = 0.853$ -- which is very close to the value I use, $f_\sigma=0.875$. This is a crude derivation neglecting the true position of energy minimum and the LJ mixing rules. However, the formula for $f_\sigma$ does not depend on the absolute position and I believe that the assumption that the minimum is relatively shifted by scaling with $f_\sigma$ holds pretty well.
